---
title: "Snow Data Assignment: Web Scraping, Functions, and Iteration"
author: "Elin Binck"
date: "2-9-2022"
output: html_document
---

```{r setup, include=FALSE}
library(rvest)
library(tidyverse)
library(lubridate)
library(readxl)
library(scales)

```


# Simple web scraping

R can read html using either rvest, xml, or xml2 packages. Here we are going to navigate to the Center for Snow and Avalance Studies  [Website](https://snowstudies.org/archived-data/) and read a table in. This table contains links to data we want to programatically download for three sites. We don't know much about these sites, but they contain incredibly rich snow, temperature, and precip data. 


## Reading an html 

### Extract CSV links from webpage

```{r}
site_url <- 'https://snowstudies.org/archived-data/'

#Read the web url
webpage <- read_html(site_url)

#See if we can extract tables and get the data that way
tables <- webpage %>%
  html_nodes('table') %>%
  magrittr::extract2(3) %>%
  html_table(fill = TRUE)
#That didn't work, so let's try a different approach

#Extract only weblinks and then the URLs!
links <- webpage %>%
  html_nodes('a') %>%
  .[grepl('24hr',.)] %>%
  html_attr('href')
```

## Data Download

### Download data in a for loop

```{r}

#Grab only the name of the file by splitting out on forward slashes
splits <- str_split_fixed(links,'/',8)

#Keep only the 8th column
dataset <- splits[,8] 

#generate a file list for where the data goes
file_names <- paste0('data/',dataset)

for(i in 1:3){
  download.file(links[i],destfile=file_names[i])
}

downloaded <- file.exists(file_names)

evaluate <- !all(downloaded)

```


### Download data in a map

```{r}

#Map version of the same for loop (downloading 3 files)
if(evaluate == T){
  map2(links[1:3],file_names[1:3],download.file)
}else{print('data already downloaded')}

```

## Data read-in 

### Read in just the snow data as a loop

```{r}
#Pattern matching to only keep certain files
snow_files <- file_names %>%
  .[!grepl('SG_24',.)] %>%
  .[!grepl('PTSP',.)]

#view(file_names)

#I don't need to repeat this in the assignment because it is just selecting for two of the three files originally downloaded


#empty_data <- list()

# snow_data <- for(i in 1:length(snow_files)){
#   empty_data[[i]] <- read_csv(snow_files[i]) %>%
#     select(Year,DOY,Sno_Height_M)
# }

#snow_data_full <- do.call('rbind',empty_data)

#summary(snow_data_full)
```


### Read in the data as a map function

```{r}

our_snow_reader <- function(file){
  name = str_split_fixed(file,'/',2)[,2] %>%
    gsub('_24hr.csv','',.)
  df <- read_csv(file) %>%
    select(Year,DOY,Sno_Height_M) %>%
    mutate(site = name)
}

snow_data_full <- map_dfr(snow_files,our_snow_reader)

summary(snow_data_full)
```


### Plot snow data

```{r}
snow_yearly <- snow_data_full %>%
  group_by(Year,site) %>%
  summarize(mean_height = mean(Sno_Height_M,na.rm=T))

ggplot(snow_yearly,aes(x=Year,y=mean_height,color=site)) + 
  geom_point() +
  ggthemes::theme_few() + 
  ggthemes::scale_color_few()
```


# Assignment:

## Extract the URLS

1. Extract the meteorological data URLs. Here we want you to use the `rvest` package to get the URLs for the `SASP forcing` and `SBSP_forcing` meteorological datasets.

```{r}
#Extract the weblinks and then the URLs
links2 <- webpage %>%
  html_nodes('a') %>%
  .[grepl('forcing',.)] %>%
  html_attr('href')

```

##Download the data

2. Download the meteorological data. Use the `download_file` and `str_split_fixed` commands to download the data and save it in your data folder. You can use a for loop or a map function. 

```{r}
#Grab only the name of the file by splitting out on forward slashes
splits2 <- str_split_fixed(links2,'/',8)

#Keep only the 8th column
dataset2 <- splits2[,8] 

#generate a file list for where the data goes
file_names2 <- paste0('data/',dataset2)

for(i in 1:2){
  download.file(links2[i],destfile=file_names2[i])
}

downloaded2 <- file.exists(file_names2)

evaluate2 <- !all(downloaded2)

```

##Read in the data

3. Write a custom function to read in the data and append a site column to the data. 

```{r}

# this code grabs the variable names from the metadata pdf file
library(pdftools)
headers <- pdf_text('https://snowstudies.org/wp-content/uploads/2022/02/Serially-Complete-Metadata-text08.pdf') %>%
  readr::read_lines(.) %>%
  trimws(.) %>%
  str_split_fixed(.,'\\.',2) %>%
  .[,2] %>%
  .[1:14] %>%
  str_trim(side = "left")


```

4. Use the `map` function to read in both meteorological files. Display a summary of your tibble.

```{r}
name = str_split_fixed(file_names2,'_',3)[,2]

our_meteorological_reader <- function(file){
  name = str_split_fixed(file,'_',3)[,2]
  df <- read.delim(file, header = F, skip = 4, sep = "") %>% 
    select(1:14) %>% 
    setNames(headers) %>% 
    mutate("site" = name) 
  }

met_data_full <- map_dfr(file_names2, our_meteorological_reader)

summary(met_data_full)

```

##Plot the annual mean temperatures

5. Make a line plot of mean temp by year by site (using the `air temp [K]` variable). Is there anything suspicious in the plot? Adjust your filtering if needed.

Two years of data prior to 2005 appeared abnormal, and were removed.  The plot shows that the Senator Beck Study Plot has mean annual temperatures that are consistently around 1.5 degrees lower than the Snow Angel Study Plot.


```{r}

#make a df with yearly average temps, rename the column since the spaces are resulting in errors

yearly_means<- met_data_full %>% 
  rename(airTempK = "air temp [K]") %>% 
    filter(year > 2005) %>% 
  group_by(year, site) %>% 
  summarise(mean_temp = mean(airTempK, na.rm = T))

#plot the data

ggplot(yearly_means, aes(x=year, y = mean_temp, color = site)) +
  geom_point()+
  theme_bw()
  
```

##Write a function to plot the mean monthly temps

6. Write a function that makes line plots of monthly average temperature at each site for a given year. Use a for loop to make these plots for 2005 to 2010. Are monthly average temperatures at the Senator Beck Study Plot ever warmer than the Snow Angel Study Plot?
Hint: https://ggplot2.tidyverse.org/reference/print.ggplot.html

Monthly average temperatures at the Senator Beck Study Plot were consistently lower than at the Snow Angel Study Plot from 2005 to 2010. 

```{r warning = FALSE}
our_meteorological_reader <- function(file){
  name = str_split_fixed(file,'_',3)[,2]
  df <- read.delim(file, header = F, skip = 4, sep = "") %>% 
    select(1:14) %>% 
    setNames(headers) %>% 
    mutate("site" = name) 
  }

monthly_means<- met_data_full %>% 
  rename(airTempK = "air temp [K]") %>% 
  group_by(month, year,site) %>% 
  summarise(mean_temp = mean(airTempK, na.rm = T))

plot_monthly_means<- function(year.no) {
  plot<-monthly_means %>% 
    filter(year == year.no) %>% 
    ggplot(aes(x = month, y = mean_temp, color = site))+
    geom_line()+
    scale_x_discrete(name = "Month", limits = c(1,2,3,4,5,6,7,8,9,10,11,12))+
    ylab("Mean Temperature (K)")+
    ggtitle(year.no)+
    ggthemes::theme_few()
    print(plot)
}

for(i in 2005:2010){
  plot_monthly_means(i)
}

```


Bonus: Make a plot of average daily precipitation by day of year (averaged across all available years). Color each site. 

Bonus #2: Use a function and for loop to create yearly plots of precipitation by day of year. Color each site. 
